---
title: "COVID-19 Analysis for Dennis Woodson"
author: "David J Jackson"
date: "August 1, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=9, fig.height=6)
library(tidyverse)
library(lubridate)
library(RSQLite)
library(scales)
library(plotly)
theme_set(theme_light())
```

## What is a Analysis?

Analysis is the process of breaking a complex topic or substance into smaller parts in order to gain a better understanding of it. The technique has been applied in the study of mathematics and logic since before Aristotle, though analysis as a formal concept is a relatively recent development.(Wikipedia)

```{r,echo=FALSE}
rm(list=ls())
db <- dbConnect(SQLite(), dbname="../COVIDDB/COVID.sqlite3")
NYT <- dbGetQuery(db, "SELECT * FROM USA ;")
USA <-  dbGetQuery(db, "SELECT * FROM ECDC  WHERE Countries ='USA';")
JHU <- dbGetQuery(db, "SELECT * FROM JHUDATA WHERE country='US' ;")
dbDisconnect(db)
 
USA$Reported <- ymd(USA$Reported)
JHU$date <- ymd(JHU$date)
NYT$date <- ymd(NYT$date)
dbDisconnect(db)
```
```{r, echo=FALSE}
JHU$DC <- JHU$confirmed - lag(JHU$confirmed)
JHU$DD <- JHU$deaths - lag(JHU$deaths)

NYT$DC <- NYT$cases - lag(NYT$cases)
NYT$DD <-NYT$deaths - lag(NYT$deaths)
```

## The Whole Shootin Match: Jan. 23,2020 - August 1, 2020
### Total Daily Cases and Deaths

```{r, echo=FALSE}
plot_1 <-USA %>% ggplot() + geom_line(aes(x=Reported,y=Cases,col="ECDPC")) +
    geom_line(data=JHU,aes(x=date,y=DC,col="John Hopkins")) +
    geom_line(data=NYT,aes(x=date,y=DC,col="NY Times")) +
  labs(title="USA Daily Cases: John Hopkins/NY Times and European CDPC")

plot_2 <- USA %>% ggplot() + geom_line(aes(x=Reported,y=Deaths,col="ECPDC")) +
    geom_line(data=JHU,aes(x=date,y=DD,col="John Hopkins")) +
    geom_line(data=NYT,aes(x=date,y=DD,col="NY Times")) +
  labs(title="UDA Daily Deaths: John Hopkins/NY Times and European CDPC")
ggplotly(plot_1)
ggplotly(plot_2)
```

### Total Cumulative Cases

```{r, echo=FALSE}
plot_3 <- NYT%>% ggplot() + geom_line(aes(x=date,y=cases,col="NY Times")) +
  geom_line(data=JHU,aes(x=date,y=confirmed,col="John Hopkins")) +
  labs(title="NYTimes/John Hopkins: US Cumulative Cases",
      subtitle = "(Jan. 23, 2020 - July 30,2020")
ggplotly(plot_3)

plot_3A <- NYT%>% ggplot() + geom_line(aes(x=date,y=deaths,col="NY Times")) +
  geom_line(data=JHU,aes(x=date,y=deaths,col="John Hopkins")) +
  labs(title="NYTimes/John Hopkins: US Cumulative deaths",
      subtitle = "(Jan. 23, 2020 - July 30,2020")
ggplotly(plot_3A)
```

### Weekly Totals for Cases and Deaths

```{r, echo=FALSE}
USA_Weekly <- USA %>% group_by(Weekly = floor_date(Reported,"week")) %>%
  summarise(across(Cases:Deaths,list(Sum = sum)))
  
JHU_Weekly <- JHU %>% group_by(Weekly = floor_date(date,"week")) %>%
  summarise(across(DC:DD,list(Sum=sum)))

NYT_Weekly <- NYT %>% group_by(Weekly = floor_date(date,"week")) %>%
  summarise(across(cases:deaths,list(Sum = sum)))
```
```{r}
plot_4 <- ggplot() + geom_col(data=JHU_Weekly,aes(x=Weekly,y=DC_Sum)) +
  labs(title="John Hopkins Weekly US Cases")
  
plot_5 <- ggplot() + geom_col(data=JHU_Weekly,aes(x=Weekly,y=DD_Sum)) +
  labs(title="John Hopkins Weekly US Deaths")      
```
```{r, echo=FALSE}
# plot_6 <- ggplot() + geom_col(data=NYT_Weekly,aes(x=Weekly,y=cases_Sum)) +
      labs(title="John Hopkins Weekly Cases")
```
```{r, echo=FALSE}
# plot_7 <- ggplot() + geom_col(data=NYT_Weekly,aes(x=Weekly,y=deaths_Sum)) +
      labs(title="NY Times Weekly Cases")
```
```{r, echo=FALSE}
ggplotly(plot_4)
ggplotly(plot_5)
# ggplotly(plot_6)
# ggplotly(plot_7)
```

##  First  Spike: In Daily U.S. Cases and Deaths April 1- 30 2020 

```{r,echo=FALSE}
A <- USA %>% filter(Reported >="2020-04-01" & Reported <="2020-04-30")
B <- JHU %>% filter(date >="2020-04-01" & date <="2020-04-30")
C <- NYT %>% filter(date >="2020-04-01" & date <="2020-04-30")

A %>% ggplot() + geom_line(aes(x=Reported,y=Cases,col="USA")) +
    geom_line(data=B,aes(x=date,y=DC,col="John Hopkins")) +
    geom_line(data=C,aes(x=date,y=DC,col="NY Times")) +
  labs(title="First Spike in U.S. Daily Cases (April 2020")

A %>% ggplot() + geom_line(aes(x=Reported,y=Deaths,col="USA")) +
    geom_line(data=B,aes(x=date,y=DD,col="John Hopkins")) +
    geom_line(data=C,aes(x=date,y=DD,col="NY Times")) +
  labs(title="First Spike in U.S.  Daily Death for  the US")
```

### Current Spike in U.S. Daily Cases and Deaths: May 1, 2020 - Present

```{r}
AA<- USA %>% filter(Reported >="2020-05-01" )
BA<- JHU %>% filter(date >="2020-05-01" )
CA <- NYT %>% filter(date >="2020-05-01")
```
```{r}
AA %>% ggplot() + geom_line(aes(x=Reported,y=Cases,col="USA")) +
    geom_line(data=BA,aes(x=date,y=DC,col="John Hopkins")) +
    geom_line(data=CA,aes(x=date,y=DC,col="NY Times")) +
  labs(title="Current  Spike in Cases: May  1, 2020 - Present")

AA %>% ggplot() + geom_line(aes(x=Reported,y=Deaths,col="USA")) +
    geom_line(data=B,aes(x=date,y=DD,col="John Hopkins")) +
    geom_line(data=C,aes(x=date,y=DD,col="NY Times")) +
  labs(title="Current Spike in Deaths: May  1, 2020 - Present")
```
